We are not the 99 percent: quantifying asphericity in the distribution of Local Group satellites

### 2017-11-26 paper/paper.tex
results form illustrisdm in the paper

\begin{figure*}
\centering
\includegraphics[width=0.30\textwidth]{scatter_random_ranked_width.pdf}
\includegraphics[width=0.30\textwidth]{scatter_ranked_width.pdf}
\includegraphics[width=0.30\textwidth]{scatter_norm_ranked_width.pdf}
\caption{Plane width characterization in the Local Group and the
  Isolated Pairs. In all panels the horizontal axis corresponds to the
  M31 or the most massive halo in the pair and the vertical axis to
  the MW or the least massive halo in the pair.
The panel on the left shows the plane width in physical units
comparing the results from observations
(stars) against the result of spherically randomizing the satellite
positions while keeping its radial distance (circles). 
The panel in the middle compares the average from the observations
(star) and the average from each one of the Isolated Pairs (circles
with error bars).
The panel on the right has the same information as the middle panel,
only that this time each point has been normalized (median substracted
and normalized by the standar deviation) to the results of its
randomization. 
The main message of this series of plots is that the MW has a
significantly thiner plane both compared to the result of its own
satellite spherical randomization (left panel) and the expectation from
simulations (middle panel). 
This low value is $2\sigma$ away from what is expected in a spherical
distribution. 
In the M31 their satellites are in agreement with the expectations
both from an spherical distribution and the results form the
simulations. 
A second conclusion is that the spherically averaged plane width
MW (seen in the point cloud in the left panel) is smaller than the
average expectation from simulations, while for M31 the spherical
average is consistent with simulations. 
\label{fig:scatter_width}}
\end{figure*}

\begin{figure*}
\centering
\includegraphics[width=0.30\textwidth]{scatter_random_ranked_ca_ratio.pdf}
\includegraphics[width=0.30\textwidth]{scatter_ranked_ca_ratio.pdf}
\includegraphics[width=0.30\textwidth]{scatter_norm_ranked_ca_ratio.pdf}
\caption{Same layout as in Figure \ref{fig:scatter_width}. 
This time for the $c/a$ axis ratio. 
The message holds in this case as for the plane width.
The MW is has a significantly low $c/a$ value compared to the
expectation from a spherical distribution and simulations. 
This low value is also $2\sigma$ away from the expactations for an
spherical distribution.
M31 is consistent both with an spherical distribution and the results
from simulations.
However, in this case the axis ratio in the spherically averaged case
is completeley consistent with the expectation from simulations.
\label{fig:scatter_ca_ratio}}
\end{figure*}

\begin{figure*}
\centering
\includegraphics[width=0.30\textwidth]{scatter_random_ranked_ba_ratio.pdf}
\includegraphics[width=0.30\textwidth]{scatter_ranked_ba_ratio.pdf}
\includegraphics[width=0.30\textwidth]{scatter_norm_ranked_ba_ratio.pdf}
\caption{Same layout as in Figure \ref{fig:scatter_width}. 
This time for the $b/a$ axis ratio. In this case both the MW and M31
are consistent with the results of a spherical distribution and the
simulations. 
\label{fig:scatter_ba_ratio}}
\end{figure*}

\begin{figure*}
\centering
\includegraphics[width=0.45\textwidth]{gaussian_model_illustris_M31.pdf}
\includegraphics[width=0.45\textwidth]{gaussian_model_illustris_MW.pdf}
\caption{Correlations between the multivariate gaussian model built on
  the normalized values for the plane width $w$, $c/a$ ratio and $b/a$ ratio. 
Left/right panel correspond to M31/MW. 
The contour levels in the 2D histograms correspond to the $1\sigma$,
$2\sigma$ and $3\sigma$ contours in two dimensions. 
The dahsed vertical lines in the histograms along the diagonal
correspond to the $1\sigma$ boundaries in one dimension.
The results for the gaussian model are built from $10^6$ point
realizations in the six-dimensional space spaned by the variables of
interest. 
The correlation matrix and the mean values are computed from the
results in the Illustris-1 simulation.
The cross indicates the LG values.
This plot clearly shows how the M31 results are well within the
expectations from simulations while MW has an unusual low value for
the plane width and the $c/a$ axis ratio.
Equivalent results from the ELVIS data are presented in Figure
\ref{fig:correlations_elvis}.
\label{fig:correlations_illustris}}
\end{figure*}

\begin{figure}
\centering
\includegraphics[width=0.47\textwidth]{expected_numbers_n_M31.pdf}
\includegraphics[width=0.47\textwidth]{expected_numbers_n_MW.pdf}
\caption{Probability distribution for the expected number of pairs
  showing the same degree of atipicality as the Local Group if drawn
  from a sample of 2000 isolated pairs. 
  The distributions correspond to results derived from Illustris-1
  and ELVIS data.
  On average between $0.05\%$ and $0.25\%$ of the isolated pairs should present
  satellite distributions as atypical as the Local Group.
\label{fig:expected_number}}
\end{figure}

>>>>

\begin{figure*}
\centering
\includegraphics[width=0.30\textwidth]{scatter_random_ranked_width.pdf}
\includegraphics[width=0.30\textwidth]{scatter_ranked_illudm_width.pdf}
\includegraphics[width=0.30\textwidth]{scatter_norm_ranked_illudm_width.pdf}
\caption{Plane width characterization in the Local Group and the
  Isolated Pairs. In all panels the horizontal axis corresponds to the
  M31 or the most massive halo in the pair and the vertical axis to
  the MW or the least massive halo in the pair.
The panel on the left shows the plane width in physical units
comparing the results from observations
(stars) against the result of spherically randomizing the satellite
positions while keeping its radial distance (circles). 
The panel in the middle compares the average from the observations
(star) and the average from each one of the Isolated Pairs (circles
with error bars).
The panel on the right has the same information as the middle panel,
only that this time each point has been normalized (median substracted
and normalized by the standar deviation) to the results of its
randomization. 
The main message of this series of plots is that the MW has a
significantly thiner plane both compared to the result of its own
satellite spherical randomization (left panel) and the expectation from
simulations (middle panel). 
This low value is $2\sigma$ away from what is expected in a spherical
distribution. 
In the M31 their satellites are in agreement with the expectations
both from an spherical distribution and the results form the
simulations. 
A second conclusion is that the spherically averaged plane width
MW (seen in the point cloud in the left panel) is smaller than the
average expectation from simulations, while for M31 the spherical
average is consistent with simulations. 
\label{fig:scatter_width}}
\end{figure*}

\begin{figure*}
\centering
\includegraphics[width=0.30\textwidth]{scatter_random_ranked_ca_ratio.pdf}
\includegraphics[width=0.30\textwidth]{scatter_ranked_illudm_ca_ratio.pdf}
\includegraphics[width=0.30\textwidth]{scatter_norm_ranked_illudm_ca_ratio.pdf}
\caption{Same layout as in Figure \ref{fig:scatter_width}. 
This time for the $c/a$ axis ratio. 
The message holds in this case as for the plane width.
The MW is has a significantly low $c/a$ value compared to the
expectation from a spherical distribution and simulations. 
This low value is also $2\sigma$ away from the expactations for an
spherical distribution.
M31 is consistent both with an spherical distribution and the results
from simulations.
However, in this case the axis ratio in the spherically averaged case
is completeley consistent with the expectation from simulations.
\label{fig:scatter_ca_ratio}}
\end{figure*}

\begin{figure*}
\centering
\includegraphics[width=0.30\textwidth]{scatter_random_ranked_ba_ratio.pdf}
\includegraphics[width=0.30\textwidth]{scatter_ranked_illudm_ba_ratio.pdf}
\includegraphics[width=0.30\textwidth]{scatter_norm_ranked_illudm_ba_ratio.pdf}
\caption{Same layout as in Figure \ref{fig:scatter_width}. 
This time for the $b/a$ axis ratio. In this case both the MW and M31
are consistent with the results of a spherical distribution and the
simulations. 
\label{fig:scatter_ba_ratio}}
\end{figure*}

\begin{figure*}
\centering
\includegraphics[width=0.45\textwidth]{gaussian_model_illustrisdm_M31.pdf}
\includegraphics[width=0.45\textwidth]{gaussian_model_illustrisdm_MW.pdf}
\caption{Correlations between the multivariate gaussian model built on
  the normalized values for the plane width $w$, $c/a$ ratio and $b/a$ ratio. 
Left/right panel correspond to M31/MW. 
The contour levels in the 2D histograms correspond to the $1\sigma$,
$2\sigma$ and $3\sigma$ contours in two dimensions. 
The dahsed vertical lines in the histograms along the diagonal
correspond to the $1\sigma$ boundaries in one dimension.
The results for the gaussian model are built from $10^6$ point
realizations in the six-dimensional space spaned by the variables of
interest. 
The correlation matrix and the mean values are computed from the
results in the Illustris-1 simulation.
The cross indicates the LG values.
This plot clearly shows how the M31 results are well within the
expectations from simulations while MW has an unusual low value for
the plane width and the $c/a$ axis ratio.
Equivalent results from the ELVIS data are presented in Figure
\ref{fig:correlations_elvis}.
\label{fig:correlations_illustris}}
\end{figure*}

\begin{figure}
\centering
\includegraphics[width=0.47\textwidth]{expected_numbers_n_M31.pdf}
\includegraphics[width=0.47\textwidth]{expected_numbers_n_MW.pdf}
\caption{Probability distribution for the expected number of M31 and
  MW halos  showing the same degree of atipicality as the Local Group if drawn
  from a sample of 10000 isolated pairs. 
  The distributions correspond to results derived from Illustris-DM, Illustris,
  and ELVIS data.
  In the most optimistic case (Illustris-DM), $1\%$ of the pairs have
  two halos with properties similar to the LG. In the case of
  Illustris this percentage drops to $0.1\%$ and $0.01\%$ for ELVIS.
\label{fig:expected_number}}
\end{figure}
